SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2021-01-22 maintex.tex
rearranged discussion subsections
minor text edits

@@

%\usepackage{epstopdf}

\begin{document}

\title{
SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs
}

\author{
Joshua~L.~Steffen\altaffilmark{1}, Hai~Fu\altaffilmark{1,2}, J.~M.~Comerford\altaffilmark{3}, Y.~Sophia~Dai\altaffilmark{4}, Shuai~Feng\altaffilmark{5,6,7}, Arran~C.~Gross\altaffilmark{1}, and Rui~Xue\altaffilmark{1}
}
\altaffiltext{1}{Department of Physics \& Astronomy, University of Iowa, Iowa City, IA 52242}
\altaffiltext{2}{Insititute for Astronomy, University of Hawaii, Honolulu, HI 96822}
\altaffiltext{3}{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, CO 80309}
\altaffiltext{4}{Chinese Academy of Sciences South America Center for Astronomy (CASSACA), 20A Datun Road, Beijing 100012, China}
\altaffiltext{5}{Department of Physics, Hebei Normal University, 20 South Erhuan Road, Shijiazhuang, 050024, China}
\altaffiltext{6}{Key Laboratory for Research in Galaxies and Cosmology, Shanghai Astronomical Observatory, Chinese Academy of Sciences, 80 Nandan Road, Shanghai 200030, China}
\altaffiltext{7}{University of the Chinese Academy of Sciences, No.19A Yuquan Road, Beijing 100049, China}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/bpt-cmd.pdf}
\caption[]{(\textbf{Left}) Color-magnitude diagram for MaNGA galaxies. \textbf{(\textbf{Center}) BPT diagram for the MaNGA galaxies. (\textbf{Right}) The WHAN diagram for the MaNGA galaxies.} The grey circles represent the whole MaNGA sample and the color scale reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. The {\it blue} diamonds represent the star forming paired galaxies (by the BPT and \ewha\ cuts), the {\it red} diamonds represent the paired galaxies which are classified as retired (by the \ewha\ cut), and the {\it green} diamonds represent the paired galaxies containing an AGN (as identified by the BPT diagnostic).}
\label{fig:cmd}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The \textbf{top} row is the absolute value of the mass ratio and the \textbf{bottom} row is the mass ratio without taking the absolute value. Taking the mass ratio without the absolute value allows us to separately study the more massive companions of pairs from less massive companions from pairs. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in \S~\ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the \textbf{169} paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

\textbf{In the top left of Figure \ref{fig:ssfr_dm}, we split difference profiles into four mass ratio bins from $|\Delta$log(M)$|$ $=$ 0.0$-$1.0. Here, $|\Delta$log(M)$|$ $\le$ 0.5, represents equal mass major mergers while $|\Delta$log(M)$|$ $\ge$ 0.5 represents minor mergers. We see that there is a clear preference for pairs in major merger to have higher sSFR enhancements in their centers, $\sim$0.3$-$0.4 dex, while minor mergers show a weaker sSFR enhancement of $\sim$0.0$-$0.2 dex. The effect can be seen more clearly with the data extracted from the central 0.5 \reff\ in the top right panel. Here the $\Delta$log(sSFR) falls linearly with wider mass ratios, reaching zero enhancement to the sSFR at $|\Delta$log(M)$|$ $=$ 1.5$-$1.75.}

@@

In \textbf{bottom left panel} of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins. \textbf{The primary companions are the pairs with $\Delta$log(M) $\ge$ 0.0 and the secondary companions are the pairs with $\Delta$log(M) $\le$ 0.0. When splitting the mass ratio like this we see that major mergers still higher levels of sSFR enhancement in comparison to minor mergers; however, we also see that the less massive companion of the pair sees a higher $\Delta$log(sSFR) in comparison to the more massive companion of a pair. In major mergers, the sSFR enhancement is $\sim$0.1 dex higher in secondary companions in comparison to primary companions and in minor mergers the sSFR enhancement is $\sim$0.2 dex higher in secondary companions in comparison to primary companions.}

@@

\textbf{We also see that the $\Delta$log(sSFR) in the outskirts ($R$ $>$ 1.5 \reff) of the pairs have a dependence on the mass ratio in the bottom right of Figure \ref{fig:ssfr_dm}. Primary companions show a zero to positive enhancement of 0.0$-$0.4 dex to the sSFR in their outskirts, while secondary companions feature a suppression to the sSFR in their outskirts of about 0.0$-$0.4 dex. Again we note that only 37\% of the MaNGA sample is designed to extend beyond 1.5 \reff; however, a similar result has been observed in the hydrodynamical simulations of \citet{Moreno:2015} and \citet{Moreno:2020}. }

@@

\section{Physical Scenario}
% General Merger-SFR enhancement
\textbf{
Our results are consistent with the idea that galaxy merger events trigger a burst of star formation in the centers of the paired galaxies. In this model, when two galaxy pairs undergo their first pericenter, the tidal torques exerted on the disks of interacting galaxies form barred structures. These barred structures funnel gases into the centers of the galaxies which triggers a burst of new star formation. Eventually, as the pairs separate from each other after the first pericenter, the burst of star formation begins to subside. This is shown in our work and many previous works in Figure \ref{fig:nuc_sep} where the sSFR enhancement is greatest for close separations and falls with wider separations. 
}

\subsection{Comparison with Previous Work}

\subsubsection{Radial Difference Profiles}

@@

In Figure \ref{fig:prof_comp} we compare the $\Delta$log(sSFR) profile and $\Delta$log(SFR) as a function of galactocentric radius between our work and previous works. \textbf{\citet{Thorp:2019} studies a sample of 36 post-mergers from the MaNGA sample. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(SFR) $=$ 0.40$-$0.45, roughly 0.2 dex higher than our paired galaxies. Between 0.4$-$1.0\reff\ the post-merger galaxies and our pairs are in closer agreement, being 0.05$-$0.1 dex higher than our paired galaxies. Beyond 1.0 \reff\ the post-merger galaxies have a higher $\Delta$log(SFR) by about 0.1$-$0.2 dex.} The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

@@

The reason for the discrepancies between our samples may be that \citet{Pan:2019} includes a set of \textbf{96 pairs which are too close to be separately de-blended by SDSS and instead were visually identified by their morphology. 82 of these pairs are post-merger galaxies. \citet{Pan:2019} shows that pairs with morphological features show higher $\Delta$log(SFR) when compared to pairs without morphological distortions. Further, \citet{Thorp:2019} showed that post-merger galaxies have $\Delta$log(SFR) profiles which are higher than our sample's profiles. From this, infer that the lower $\Delta$log(sSFR) seen in our sample is due to our sample's lack of post-merger galaxies. }

\subsection{Dependency on Merger Parameters}

@@

>>>>

%\usepackage{CJK}
%\setCJKmainfont{SimSun}

@@

\begin{document}
\title{
SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs
}

\author{
Joshua~L.~Steffen\altaffilmark{1}, Hai~Fu\altaffilmark{1,2}, J.~M.~Comerford\altaffilmark{3}, Y.~Sophia~Dai \altaffilmark{4}, Shuai~Feng\altaffilmark{5,6,7}, Arran~C.~Gross\altaffilmark{1}, and Rui~Xue\altaffilmark{1}
}

\altaffiltext{1}{Department of Physics \& Astronomy, University of Iowa, Iowa City, IA 52242}
\altaffiltext{2}{Insititute for Astronomy, University of Hawaii, Honolulu, HI 96822}
\altaffiltext{3}{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, CO 80309}
\altaffiltext{4}{Chinese Academy of Sciences South America Center for Astronomy (CASSACA), 20A Datun Road, Beijing 100012, China}
\altaffiltext{5}{Department of Physics, Hebei Normal University, 20 South Erhuan Road, Shijiazhuang, 050024, China}
\altaffiltext{6}{Key Laboratory for Research in Galaxies and Cosmology, Shanghai Astronomical Observatory, Chinese Academy of Sciences, 80 Nandan Road, Shanghai 200030, China}
\altaffiltext{7}{University of the Chinese Academy of Sciences, No.19A Yuquan Road, Beijing 100049, China}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/bpt-cmd.pdf}
\caption[]{(\textbf{Left}) Color-magnitude diagram for MaNGA galaxies. \textbf{(\textbf{Center}) BPT diagram for the MaNGA galaxies. (\textbf{Right}) The WHAN diagram for the MaNGA galaxies.} The grey circles represent the whole MaNGA sample and the color scale reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. The {\it purple} diamonds represent the paired galaxies which are classified as retired (by the \ewha\ cut), the {\it blue} diamonds represent the star forming paired galaxies (by the BPT diagram), the {\it green} diamonds represent the composite star forming-AGN paired galaxies (by the BPT diagram), and the {\it Red} diamonds represent the paired galaxies containing an AGN (by the BPT diagram).}
\label{fig:cmd}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The \textbf{top} row is the absolute value of the mass ratio and the \textbf{bottom} row is the mass ratio without taking the absolute value. Taking the mass ratio without the absolute value allows us to separately study the more massive companions of pairs from the less massive companions from pairs. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in \S~\ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the \textbf{169} paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

\textbf{In the top left of Figure \ref{fig:ssfr_dm}, we split the $\Delta$log(sSFR) profiles into four mass ratio bins from $|\Delta$log(M)$|$ $=$ 0.0$-$1.0. Here, $|\Delta$log(M)$|$ $\le$ 0.5, represents equal mass major mergers while $|\Delta$log(M)$|$ $\ge$ 0.5 represents minor mergers. We see that there is a clear preference for pairs in major merger to have higher sSFR enhancements in their centers, $\sim$0.3$-$0.4 dex, while minor mergers show a weaker sSFR enhancement of $\sim$0.0$-$0.2 dex. The effect can be seen more clearly with the data extracted from the central 0.5 \reff\ in the top right panel. Here the $\Delta$log(sSFR) falls linearly with wider mass ratios, reaching zero enhancement to the sSFR at $|\Delta$log(M)$|$ $=$ 1.5$-$1.75.}

@@

In \textbf{bottom left panel} of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins. \textbf{The primary companions are the pairs with $\Delta$log(M) $\ge$ 0.0 and the secondary companions are the pairs with $\Delta$log(M) $\le$ 0.0. When splitting the mass ratio like this we see that major mergers still higher levels of sSFR enhancement in comparison to minor mergers; however, we also see that the less massive companion of the pair sees a higher $\Delta$log(sSFR) in comparison to the more massive companion of a pair. In major mergers, the sSFR enhancement is $\sim$0.1 dex higher in secondary companions in comparison to primary companions. In minor mergers the sSFR enhancement is $\sim$0.2 dex higher in secondary companions in comparison to primary companions.}

@@

\textbf{We also see that the $\Delta$log(sSFR) in the outskirts ($R$ $>$ 1.5 \reff) of the pairs have a dependence on the mass ratio in the bottom right of Figure \ref{fig:ssfr_dm}. Primary companions show a positive enhancement of 0.0$-$0.4 dex to the sSFR in their outskirts, while secondary companions feature a suppression to the sSFR in their outskirts of about 0.0$-$0.3 dex. Again we note that only 37\% of the MaNGA sample is designed to extend beyond 1.5 \reff; however, a similar result has been observed in the hydrodynamical simulations of \citet{Moreno:2015} and \citet{Moreno:2020}. }

@@

\subsection{Radial Profiles of Enhancement}

@@

In Figure \ref{fig:prof_comp} we compare the $\Delta$log(sSFR) profile and $\Delta$log(SFR) as a function of galactocentric radius between our work and previous works using the MaNGA data. \textbf{\citet{Thorp:2019} studies a sample of 36 post-mergers from the MaNGA sample. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(SFR) $=$ 0.40$-$0.45, roughly 0.2 dex higher than our paired galaxies. Between 0.4$-$1.0\reff\ the post-merger galaxies and our pairs are in closer agreement, being 0.05$-$0.1 dex higher than our paired galaxies. Beyond 1.0 \reff\ the post-merger galaxies have a higher $\Delta$log(SFR) by about 0.1$-$0.2 dex.} The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

@@

The reason for the discrepancies between our samples may be that \citet{Pan:2019} includes a set of \textbf{96 pairs which are too close to be separately de-blended by SDSS and instead were visually identified by their morphology. 82 of these pairs are post-merger galaxies. \citet{Pan:2019} showed that pairs with disturbed morphology have higher $\Delta$log(SFR) when compared to pairs without morphological distortions. Further, \citet{Thorp:2019} showed that post-merger galaxies have $\Delta$log(SFR) profiles which are elevated over our sample's profiles. From this, infer that the lower $\Delta$log(sSFR) seen in our sample is due to our sample's lack of post-merger galaxies. }

\subsection{Central Enhancement vs. Merger Parameters}

@@

\subsection{Comparison with Simulations}

% General Merger-SFR enhancement
\textbf{
Our results are consistent with the idea that galaxy merger events trigger a burst of star formation in the centers of the paired galaxies. In this model, when two galaxy pairs undergo their first pericenter, the tidal torques exerted on the disks of interacting galaxies form barred structures. These barred structures funnel gases into the centers of the galaxies which triggers a burst of new star formation. Eventually, as the pairs separate from each other after the first pericenter, the burst of star formation begins to subside. This is shown in our work and many previous works in Figure \ref{fig:nuc_sep} where the sSFR enhancement is greatest for close separations and falls with wider separations. 
}
